Extracting Angular Observables without a Likelihood\\and Applications to Rare Decays

### 2014-08-04 draft.tex
More work on sec. IV, comments on Fred's points

\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[svgnames]{xcolor}
\usepackage{hepparticles}
\usepackage{hepnicenames}
\usepackage{hepunits}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{epstopdf}

%% Shortcuts %%
\newcommand{\aver}[1]{\langle #1 \rangle}
\newcommand{\dual}[1]{\tilde{#1}}
\newcommand{\est}[1]{\widehat{#1}}
\newcommand{\ie}{\textit{i.e.}}
\newcommand{\nuvec}{\vec{\nu}}
\newcommand{\order}[1]{\mathcal{O}\left({#1}\right)}
\newcommand{\refapp}[1]{app.~(\ref{app:#1})}
\newcommand{\refeq}[1]{eq.~(\ref{eq:#1})}
\newcommand{\refsec}[1]{sec.~(\ref{sec:#1})}
\newcommand{\reftab}[1]{tab.~(\ref{tab:#1})}
\newcommand{\rmdx}[1]{\mbox{d} #1 \,} % differential
\newcommand{\thvec}{\vec{\vartheta}}
\let\oldtheta\theta
\renewcommand{\theta}{\vartheta}
\let\eps\varepsilon
\newcommand{\wwhat}[1]{\widehat{\widehat{#1}}}

@@

\captionsetup{compatibility=false}

@@

The initial motivation for studying what we wish to call the \emph{Method of Moments}
is the determination of angular observables in the rare decay FCNC-mediated decay
$\bar{B}\to \bar{K}^*(\to \bar{K}\pi)\ell^+\ell^-$. However, we emphasize that
the method we describe in this letter is general, and it applies to arbitrary
decay or scattering processes. We find a previous work \cite{Dighe:1998vk}
advocates this method chiefly for the determination of angular observables in non-leptonic
$B$ decays, but also mentions the applicability to semileptonic decays. We improve
upon this previous work by studying correlations between the angular observables,
and sources of systematic uncertainties. Moreover, we will show that the method
at hand has two benefits over the usual approach based on likelihood fits.
\begin{enumerate}
    \item Likelihood fits do not always converge. \fred{Why? Mismodelling? Expand!}
    \item Using the method at hand, the joint probability distribution of the
        angular observables converges towards a multivariate normal distribution.
        This allows easy transfer of information from the experiment to interest
        theorists.
\end{enumerate}

@@

The central tenet of MC integration is the fact that the expectation value $E_P[g]$ of some function
$g(x)$ under the probability density $P(x)$,
\begin{equation}
    E_P[g] \equiv \int \rmdx{x} P(x) g(x)
\end{equation}
can be approximated \cite{MCSM-2004} by the consistent and unbiased
estimator $\est{E_P[g]}$~\cite[sec. 8.2]{James:2006zz}
\begin{equation}
    \label{eq:mc-id}
    E_P[g] \to \widehat{E_P[g]} \equiv \frac{1}{K} \sum_{k=1}^K g(x^{(k)}) \,,\,    x^{(k)} \sim P
\end{equation}
due to the strong law of large numbers for $K \to \infty$ assuming
that the variates $x^{(k)}$, $k = 1, \dots, K$ are distributed as
$P$.
Throughout this letter we denote all MC estimators with a wide hat.\\

Application of \refeq{mc-id} then yields
\begin{equation}
    \langle S_i\rangle \to \widehat{\langle S_i\rangle} = \frac{1}{K} \sum_{k=1}^{K} \dual{f}_i(x^{(k)})\,.
\end{equation}
It is often of interest to obtain the $\nuvec$- observables integrated
over certain ranges of $\nuvec$; i.e., binned measurements of $\langle
S_i\rangle$.
Therefore, we introduce the bin-integrated quantity
\begin{align}
    \langle S_i\rangle_{\vec{a},\vec{b}}
    & = \int_{\vec{a}}^{\vec{b}} \rmdx{^M \nu} S_i(\nuvec)\\
    & = \int_{\vec{a}}^{\vec{b}} \rmdx{^M \nu} \left[\int_{\Omega} \rmdx{^N\theta} P(\nuvec,\thvec) \dual{f}_i(\thvec)  \right]\\
    & = \int \rmdx{^M \nu} \left[\int_{\Omega} \rmdx{^N\theta}  P(\nuvec,\thvec) \dual{f}_i(\thvec)
        \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,
        \right],
\end{align}
where the argument of the indicator function $ \mathbf{1}(\vec{a} \le
\nuvec \le \vec{b})$ is to be interpreted componentwise.
Application of \refeq{mc-id} immediately yields
\begin{equation}
    \label{eq:bin-importance}
    \widehat{\langle S_i\rangle_{\vec{a},\vec{b}}}
    = \frac{1}{K} \sum_{k=1}^{K} \dual{f}_i(x^{(k)})         \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,.
\end{equation}
This reduces the estimation of $\aver{S_i}_{\vec{a},\vec{b}}$ to a
(weighted) counting experiment. In the limit $N \to \infty$, the
central limit theorem (CLT) implies that the random vector
\begin{equation}
  \label{eq:angular-obs-vec}
  \widehat{\vec{S}_{\vec{a},\vec{b}}} \equiv (\aver{S_1}_{\vec{a},\vec{b}}\,, \dots, \aver{S_i}_{\vec{a},\vec{b}}\,,
  \dots)
\end{equation}
 follows a multivariate Gaussian distribution whose mean vector
is approximated by \refeq{bin-importance} and whose covariance is
estimated as
\begin{equation}
    \cov[\aver{S_i},\aver{S_j}]_{\vec{a},\vec{b}} \to \est{\cov}[\aver{S}_i, \aver{S}_j]_{\vec{a},\vec{b}}
        = \frac{1}{K - 1} \sum_{k=1}^{K} \big[\dual{f}_i(x^{(k)}) - \est{\langle S_i\rangle_{\vec{a},\vec{b}}}\big]\,\big[\dual{f}_j(x^{(k)}) - \widehat{\langle S_j\rangle_{\vec{a},\vec{b}}}\big]
        \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,.
\end{equation}
In our physics applications, the parameter space is compact and each
$\dual{f}_i$ is bounded. Hence the requisites for the most basic
version of the CLT to hold --- finite mean and covariance of
$\dual{f}_i$ --- are automatically satisfied.  We would like to
emphasize that the sample covariance should \fred{``Should'' as in it
  does in practice? Reference toy studies below} rapidly converge
towards the true covariance matrix.  We propose to declare convergence
\fred{I'd prefer: check that skewness and kurtosis are sufficiently
  small} once estimators for skewness and kurtosis are sufficiently
small,
\begin{equation}
    \est{\operatorname{Skew}}[\aver{S_i},\aver{S_j}] \ll 1\,\,\text{ and }\,\,\est{\kurt}[\aver{S_i},\aver{S_j}] \ll 1\,.
\end{equation}

@@

@@

So far, we assumed that our underlying physical process was accurately described by our PDF $P$.
However, in several interesting processes we might only have an approximative result for $P$.
In this section, we take the interesting class of four-body decays $B\to P_1 P_2 \ell_1 \bar\ell_2$
as an example\footnote{This includes the rare $b\to s$ mediated $B$ decay $B \to K\pi\ell^+\ell^-$, and
the $V_{ub}$ suppressed decay $B\to \pi\pi\ell^+\bar\nu_\ell$. For both examples the PDF $P$ is known in
the small-width approximation and when assuming a pure $P$-wave, resonant final state. An extension to
$S-P$ interference has been studied for $B\to K\pi\ell^+\ell^-$ \cite{Blake:2012mb,Becirevic:2011bp},
and for $\bar{B}\to \pi\pi\ell^-\bar\nu_\ell$ \cite{Faller:2013dwa}. For a first study of $S$, $P$
and $D$ interference, see \cite{Das:2014sra}}.%
The PDFs for these decays are usually expressed in terms of one or a few partial waves
of the dimeson system. However, the angular momentum of the dimeson system is unbounded from above.
\\

For the selected class of decays, we face the problem that $P$ has a given dependence on the dilepton
helicity angle $\theta_1$ and the azimuthal angle $\theta_3$. (See also \refapp{btokpill} for details
on the angular distribution.) However, on the level of decay amplitudes the dimeson system can have an arbitrary
large total angular momentum $j$; only its third component is restricted to $m = -1,0,+1$.\\

We advocate here that is a sensible procedure to perform an expansion in terms of Legendre polynomials $p_{j}^{(|m|)}$
with respect to the remaining angle $\theta_2$. Using the observables $S_k(\vec{\nu})$, $k=1,\dots,9$, with $\vec{\nu}=(q^2, k^2)$
and as defined in \refapp{partial-waves}, the expansion reads
\begin{equation}
    S_{k}(\vec{\nu},\cos\theta_2) \equiv \sum_{j} \frac{1}{n_{j,|m|}} S_{k,j}(\vec{\nu}) p_{j}^{(|m|)}(\cos\theta_2)\sin\theta_2\,,
\end{equation}
where the the normalization factor $n_{j,|m|}$ is defined in \refeq{legendre-scalar-product}, and we have
\begin{equation}
    |m| = \begin{cases}
        0\,, & k = 1,2,6\,\\
        1\,, & k = 3,4,5,7,8,9\,,
    \end{cases}\,.
\end{equation}
The angular observables $S_{k,j}$ defined as coefficients of this expansion
have the merit of a well-defined total angular momentum, and thus are physically discriminable.
As a consequence of the orthogonality of the Legendre polynomials, any mismodelling (or rather, lack of modelling of higher
partial-wave observables) \emph{does not affect} the method of moments as discussed in the previous section.\\

@@

We wish to emphasize, however, that detector acceptance effects systematically affect the expansion for any basis
of angular functions. The means to cope with this problem are discussed in the following subsection.

\subsection{Detector Acceptance Effects}
\label{sec:systematics:acceptance}

\paragraph{Recipe to determine the acceptance effects}

The determination of detector acceptance is generally a difficult
task. Here, we will show a systematic method to estimate its effects
based on both the method of moments, and MC simulations of the
detector.

\fred{In the ideal case, one would have an explicit probabilistic
  model of the detector acceptance effects and could thus write down
  the full forward model of the measured events arise. In practice,
  this would require much more CPU time than available, and one is
  therefore forced to simplify the model. The standard approach is to
  generate \emph{true} particle events $\lbrace x_t^{(k)}\rbrace =
  \lbrace(\vec{\nu}^{(k)}_t,\vec\theta^{(k)}_t)\rbrace$ from a PDF
  assumed to describe the bare physical process, and to propagate
  those particles through a detailed simulation of the detector. The
  observable traces that the particles leave in the detector are fed
  into reconstruction algorithms resulting in the \emph{detector}
  events $\lbrace x^{(k)}_d\rbrace = \lbrace(\vec{\nu}^{(k)}_d,
  \vec\theta^{(k)}_d)\rbrace$.

[Where does the reconstruction uncertainty enter? Below we just
  account for systematic shifts but our estimate of the uncertainty of
  one $S_k$ would be only the statistical uncertainty from the finite
  sample size. How about the extra correlation between different $S_k$
  that is induced?] }\\

As before, we assume a PDF $P$ can be decomposed into $N$ parts with angular observables $\lbrace S_k\rbrace$,
\begin{equation}
    P \equiv P(\vec\nu,\vec\theta | \lbrace S_k \rbrace)\,.
\end{equation}
We denote the estimator of an
observable $S_k$ as $\est{S_k}$, which is affected by the detector acceptance. We assume now that there is a linear relation
between the true angular observables $\lbrace S_k^\text{true}\rbrace$, and the estimators $\lbrace \est{S_k}\rbrace$,
\begin{equation}
    \label{eq:linear-assumption}
    \est{S_i} = \sum_{i = 1}^N M_{i j} S_j^\text{true}\,.
\end{equation}
This assumption is based on a smooth angular acceptance function. While this will not be true for the laboratory frame,
smoothness may be recovered in boosted frame, e.g. the rest frame of the decaying particle.
In that case, the matrix $M_{ij}$ can be fully determined through the following steps.
\begin{enumerate}
    \item For a fixed $j$, let $S_j^\text{true} = 1$, and $S_k^\text{true} = 0\,\forall k\neq j$.
    \item Draw simulated events $\lbrace x_t^{(k)}\rbrace = \lbrace(\vec{\nu}^{(k)}_t,\vec\theta^{(k)}_t)\rbrace$ (dubbed ``true'') from the PDF $P_j$,
        \begin{equation}
            P_j \equiv P(\vec\nu,\vec\theta | \lbrace S_j = 1, S_{k\neq j} = 0\rbrace)\,.
        \end{equation}
    \item Use the true events as input for the MC simulation of the detector, in order to obtain the detector
        events $\lbrace x^{(k)}_d\rbrace = \lbrace(\vec{\nu}^{(k)}_d,\vec\theta^{(k)}_d)\rbrace$.
    \item Use the method of moments to determine the uncorrected observables $\lbrace \est{S_k} \rbrace$, based on the detector
        events $\lbrace x^{(k)}_d\rbrace$.
    \item By virtue of \refeq{linear-assumption} and step 1, the $j$-th column of $M$ has been determined via
        \begin{equation}
            M_{ij} = \est{S_i}\,.
        \end{equation}
        \fred{I wonder if a small sample size also leads to artificial
          nonzero off-diagonal elements of $M$}
    \item Repeat the above steps for all values of $j$ to determine all columns of $M$.
\end{enumerate}

@@

We emphasize that this methods hinges on the assumption in \refeq{linear-assumption}, We thus suggest to explicitly
test the assumption. One way of testing involves repeating the previous procedure
with $S_j^\text{true} = 1/x$. If the lineary assumption holds, one should find
\begin{equation}
    M_{ij} = x \est{S_i}\,,
\end{equation}
for arbitrary values of $x > 1$.

@@

In order to study the performance of the proposed method, we simulate measurements for one uni-angular
and one tri-angular decay distribution for various numbers of events. As inputs we use the SM predictions
for the decays $B\to K\ell^+\ell^-$ and $B\to K^*(\to K\pi)\ell^+\ell^-$, respectively. We cover the
$q^2$ ranges $1\GeV^2$ to $6\GeV^2$, and $15\GeV^2$ to to the respective kinematic endpoint, with two
bin widths: $1\GeV^2$ and $0.5\GeV^2$. This is to ensure that a wide spectrum of possible values for
the angular observables. We find:
\begin{itemize}
    \item In all studied cases we observed not a single bias in the pull distribution of the observable.
    \item The error $\sigma_k(N)$ of an angular observable $S_k$, evaluate for $N$ events is very well
        fitted by
        \begin{equation}
            \sigma_k(N) = \frac{\sigma_k(1)}{\sqrt{N}}
        \end{equation}
        with $\sigma_k(1) = \order{1}$.
    \item \todo{Plot Skewness and Kurtosis for $S_5$ and $S_7$ as functions of $N$.}
    \item \todo{Perform Mardia's Skewness and Kurtosis tests.}
    \item \dots
\end{itemize}
We find that all of the above results show consistenly convergence of the distribution for the angular observables
towards a multivariate normal distribution.

\begin{figure}[t]
        \centering
        \begin{subfigure}[b]{0.45\textwidth}
                \includegraphics[width=0.8\textwidth]{figs/Q2_5_6_S5.png}
        \end{subfigure}
        \begin{subfigure}[b]{0.45\textwidth}
                \includegraphics[width=0.8\textwidth]{figs/Q2_5_6_S7.png}
        \end{subfigure}
        \caption{Uncertainty of the angular observables $S_5$ and $S_7$, extracted from simulated events of the decay $B\to K^*\ell^+\ell^-$ and as a function of the number of simulated events $N$. The red curve represents a fit to the function $\sigma_k(N) = \frac{\sigma_k(1)}{\sqrt{N}}$.}
        \label{fig:errors}
\end{figure}

>>>>

\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[svgnames]{xcolor}
\usepackage{hepparticles}
\usepackage{hepnicenames}
\usepackage{hepunits}
\usepackage{graphicx}
\usepackage{epstopdf}

%% Shortcuts %%
\newcommand{\aver}[1]{\langle #1 \rangle}
\newcommand{\dual}[1]{\tilde{#1}}
\newcommand{\est}[1]{\widehat{#1}}
\newcommand{\ie}{\textit{i.e.}}
\newcommand{\nuvec}{\vec{\nu}}
\newcommand{\order}[1]{\mathcal{O}\left({#1}\right)}
\newcommand{\refapp}[1]{app.~(\ref{app:#1})}
\newcommand{\refeq}[1]{eq.~(\ref{eq:#1})}
\newcommand{\reffig}[1]{fig.~(\ref{fig:#1})}
\newcommand{\refsec}[1]{sec.~(\ref{sec:#1})}
\newcommand{\reftab}[1]{tab.~(\ref{tab:#1})}
\newcommand{\rmdx}[1]{\mbox{d} #1 \,} % differential
\newcommand{\thvec}{\vec{\vartheta}}
\let\oldtheta\theta
\renewcommand{\theta}{\vartheta}
\let\eps\varepsilon
\newcommand{\wwhat}[1]{\widehat{\widehat{#1}}}

@@

@@

The initial motivation for studying what we wish to call the \emph{Method of Moments}
is the determination of angular observables in the rare decay FCNC-mediated decay
$\bar{B}\to \bar{K}^*(\to \bar{K}\pi)\ell^+\ell^-$. However, we emphasize that
the method we describe in this letter is general, and it applies to arbitrary
decay or scattering processes. We find a previous work \cite{Dighe:1998vk}
advocates this method chiefly for the determination of angular observables in non-leptonic
$B$ decays, but also mentions the applicability to semileptonic decays. We improve
upon this previous work by studying correlations between the angular observables,
and sources of systematic uncertainties. Moreover, we will show that the method
at hand has two benefits over the usual approach based on likelihood fits. To wit:
\begin{enumerate}
    \item Likelihood fits have convergence problems for a small number of
        events, and can require reparametrizations and/or approximations
        for a successful fit to the signal PDF, see e.g. \cite{Egede:2008uy}.
        The method of moments has no such convergence problems.
    \item The results of likelihood fits suffer from systematic uncertainties
        and systematic bias (see e.g. \cite{Egede:2008uy}), especially in but not limited
        to cases in which the PDFs do not faithfully model the underlying physics. We will
        show that the method of moments does not suffer a very particular type of
        mismodelling that arises from introducing a cutoff in a partial wave expansion.
    \item Using the method of moments, the joint probability distribution of the
        angular observables converges towards a multivariate normal distribution.
        This allows an easy transfer of correlation information from the experiments to
        interested theorists.
\end{enumerate}

@@

The central tenet of MC integration is the fact that the expectation value $E_P[g]$ of some function
$g(x)$ under the probability density $P(x)$,
\begin{equation}
    E_P[g] \equiv \int \rmdx{x} P(x) g(x)
\end{equation}
can be approximated \cite{MCSM-2004} by the consistent and unbiased
estimator $\est{E_P[g]}$~\cite[sec. 8.2]{James:2006zz}
\begin{equation}
    \label{eq:mc-id}
    E_P[g] \to \widehat{E_P[g]} \equiv \frac{1}{K} \sum_{k=1}^K g(x^{(k)}) \,,\,    x^{(k)} \sim P
\end{equation}
due to the strong law of large numbers for $K \to \infty$, assuming
that the variates $x^{(k)}$, $k = 1, \dots, K$ are distributed as
$P$.
Throughout this letter we denote all MC estimators with a wide hat.\\

Application of \refeq{mc-id} then yields
\begin{equation}
    \langle S_i\rangle \to \widehat{\langle S_i\rangle} = \frac{1}{K} \sum_{k=1}^{K} \dual{f}_i(x^{(k)})\,.
\end{equation}
It is often of interest to obtain the $\nuvec$- observables integrated
over certain ranges of $\nuvec$; i.e., binned measurements of $\langle
S_i\rangle$.
Therefore, we introduce the bin-integrated quantity
\begin{align}
    \langle S_i\rangle_{\vec{a},\vec{b}}
    & = \int_{\vec{a}}^{\vec{b}} \rmdx{^M \nu} S_i(\nuvec)\\
    & = \int_{\vec{a}}^{\vec{b}} \rmdx{^M \nu} \left[\int_{\Omega} \rmdx{^N\theta} P(\nuvec,\thvec) \dual{f}_i(\thvec)  \right]\\
    & = \int \rmdx{^M \nu} \left[\int_{\Omega} \rmdx{^N\theta}  P(\nuvec,\thvec) \dual{f}_i(\thvec)
        \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,
        \right],
\end{align}
where the argument of the indicator function $ \mathbf{1}(\vec{a} \le
\nuvec \le \vec{b})$ is to be interpreted componentwise.
Application of \refeq{mc-id} immediately yields
\begin{equation}
    \label{eq:bin-importance}
    \widehat{\langle S_i\rangle_{\vec{a},\vec{b}}}
    = \frac{1}{K} \sum_{k=1}^{K} \dual{f}_i(x^{(k)})         \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,.
\end{equation}
This reduces the estimation of $\aver{S_i}_{\vec{a},\vec{b}}$ to a
(weighted) counting experiment. In the limit $N \to \infty$, the
central limit theorem (CLT) implies that the random vector
\begin{equation}
  \label{eq:angular-obs-vec}
  \widehat{\vec{S}_{\vec{a},\vec{b}}} \equiv (\aver{S_1}_{\vec{a},\vec{b}}\,, \dots, \aver{S_i}_{\vec{a},\vec{b}}\,,
  \dots)
\end{equation}
 follows a multivariate Gaussian distribution whose mean vector
is approximated by \refeq{bin-importance} and whose covariance is
estimated as
\begin{equation}
    \cov[\aver{S_i},\aver{S_j}]_{\vec{a},\vec{b}} \to \est{\cov}[\aver{S}_i, \aver{S}_j]_{\vec{a},\vec{b}}
        = \frac{1}{K - 1} \sum_{k=1}^{K} \big[\dual{f}_i(x^{(k)}) - \est{\langle S_i\rangle_{\vec{a},\vec{b}}}\big]\,\big[\dual{f}_j(x^{(k)}) - \widehat{\langle S_j\rangle_{\vec{a},\vec{b}}}\big]
        \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,.
\end{equation}
In our physics applications, the parameter space is compact and each
$\dual{f}_i$ is bounded. Hence the requisites for the most basic
version of the CLT to hold --- finite mean and covariance of
$\dual{f}_i$ --- are automatically satisfied.  We would like to
emphasize that the sample covariance should \fred{``Should'' as in it
  does in practice? Reference toy studies below} rapidly converge
towards the true covariance matrix.  We propose to declare convergence
\fred{I'd prefer: check that skewness and kurtosis are sufficiently
  small} once estimators for skewness and kurtosis are sufficiently
small,

\begin{equation}
    \est{\operatorname{Skew}}[\aver{S_i},\aver{S_j}] \leq B_S\,\,\text{ and }\,\,\est{\kurt}[\aver{S_i},\aver{S_j}] \leq B_K\,.
\end{equation}
where the limits $B_S$ and $B_K$ should be determined from process-specific, dedicated toy studies.\\

@@

  \danny{Two comments:
      \begin{itemize}
      \item We should tune down the ``data privacy'' line of thought. LHCb
      actively works on providing the full likelihood to theorists.
      Rather, we should emphasize that providing the likelihood
      as obtained from the ML-fit is technically much harder (think: profile likelihood,
      since the theorists are not interested in control channels, \dots)
      \item The constraints as described here are \emph{theoretical}
      in nature! If one changes the PDF (e.g. in $B\to K^*\ell^+\ell^-$,
      higher partial waves will have an impact on the shape of the
      \emph{physical region}. Thus, imposing these constraints should
      not be done on the side of the experiment.
      \end{itemize}
  }

@@

So far, we assumed that our underlying physical process was accurately described by our PDF $P$.
However, in several interesting processes we might only have an approximative result for $P$.
In this section we focus on one particular class of mismodelling of the signal PDF: the angular
momentum-cutoff in partial wave expansions. This mismodelling potentially affects a large
number of decays and scattering processes. For the sake of clarity we take the interesting class
of four-body decays $B\to P_1 P_2 \ell_1 \bar\ell_2$ as an example\footnote{This includes the rare $b\to s$ mediated $B$ decay $B \to K\pi\ell^+\ell^-$, and
the $V_{ub}$ suppressed decay $B\to \pi\pi\ell^+\bar\nu_\ell$. For both examples the PDF $P$ is known in
the small-width approximation and when assuming a pure $P$-wave, resonant final state. An extension to
$S-P$ interference has been studied for $B\to K\pi\ell^+\ell^-$ \cite{Blake:2012mb,Becirevic:2011bp},
and for $\bar{B}\to \pi\pi\ell^-\bar\nu_\ell$ \cite{Faller:2013dwa}. For a first study of $S$, $P$
and $D$ interference, see \cite{Das:2014sra}}.\\

Within existing analyses, the PDFs for these decays are usually expressed in terms of one or a few
partial waves of the dimeson system. However, the angular momentum of the dimeson system is unbounded
from above, and gives rise to an infinite set of angular observables.\\

For the selected class of decays, we can describe that problem as follows: The PDF $P$ has a fixed dependence on the dilepton
helicity angle $\theta_1$ and the azimuthal angle $\theta_3$. (See also \refapp{btokpill} for details
on the angular distribution.) However, on the level of decay amplitudes the dimeson system can have an arbitrarily
large total angular momentum $j$; only its third component is restricted to $m = -1,0,+1$.
We advocate here that is a sensible procedure to perform an expansion in terms of Legendre polynomials $p_{j}^{(|m|)}$
with respect to the remaining angle $\theta_2$. Using the observables $S_k(\vec{\nu})$, $k=1,\dots,9$, with $\vec{\nu}=(q^2, k^2)$
and as defined in \refapp{partial-waves}, the expansion reads
\begin{equation}
    S_{k}(\vec{\nu},\cos\theta_2) \equiv \sum_{j} \frac{1}{n_{j,|m|}} S_{k,j}(\vec{\nu}) p_{j}^{(|m|)}(\cos\theta_2)\sin\theta_2\,,
\end{equation}
where the the normalization factor $n_{j,|m|}$ is defined in \refeq{legendre-scalar-product}. Within our
example we have (cf. also \refapp{btokpill})
\begin{equation}
    |m| = \begin{cases}
        0\,, & k = 1,2,6\,\\
        1\,, & k = 3,4,5,7,8,9\,,
    \end{cases}\,.
\end{equation}
The angular observables $S_{k,j}$ defined as coefficients of this expansion in Legendre polynomials
have the merit of a well-defined total angular momentum, and thus are physically discriminable.
As a consequence of the orthogonality of the Legendre polynomials, any mismodelling (or rather, lack of modelling of higher
partial-wave observables) \emph{does not affect} the method of moments as discussed in the previous section. To wit,
ignoring further (orthogonal) terms in the PDF cannot affect the mean vector $(\dots, S_k, \dots)$ nor its variance.\\

@@

We wish to emphasize, however, that detector acceptance effects systematically affect the expansion for any basis
of angular functions, including the one suggested in this section. The means to cope with this problem are discussed
in the following subsection.

\subsection{Recipe for the Determination of Detector Acceptance Effects}
\label{sec:systematics:acceptance}

The determination of a detector's angular acceptance is generally a difficult
task. Here, we will show a systematic method to estimate its effects
based on both the method of moments, and MC simulations of the
detector.\\

In the ideal case, one would have an explicit probabilistic
model of the detector acceptance effects and could thus write down
the full forward model from which the measured events arise. In practice,
this would require much more CPU time than available, and one is
therefore forced to simplify the model. The standard approach is to
generate \emph{true} particle events $\lbrace x_t^{(k)}\rbrace =
\lbrace(\vec{\nu}^{(k)}_t,\vec\theta^{(k)}_t)\rbrace$ from a PDF
assumed to describe the bare physical process, and to propagate
those particles through a detailed simulation of the detector. The
observable traces that the particles leave in the detector are fed
into reconstruction algorithms resulting in the \emph{detector}
events $\lbrace x^{(k)}_d\rbrace = \lbrace(\vec{\nu}^{(k)}_d,
\vec\theta^{(k)}_d)\rbrace$.\\

As before, we assume the PDF $P$ can be decomposed into $N$ parts with angular observables $\lbrace S_k\rbrace$,
\begin{equation}
    P \equiv P(\vec\nu,\vec\theta | \lbrace S_k \rbrace)\,.
\end{equation}
We denote the estimator of an
observable $S_k$ as $\est{S_k}$, which is affected by the detector acceptance. We assume now that there is a linear relation
between the true angular observables $\lbrace S_k^\text{true}\rbrace$, and the estimators $\lbrace \est{S_k}\rbrace$,
\begin{equation}
    \label{eq:linear-assumption}
    \est{S_i} = \sum_{j = 1}^N M_{i j} S_j^\text{true}\,.
\end{equation}
This assumption is based on a smooth angular acceptance function. While this will not be true for the laboratory frame,
smoothness may be recovered in a boosted frame, e.g. the rest frame of the decaying particle.
In that case, the matrix $M_{ij}$ can be fully determined through the following steps.
\begin{enumerate}
    \item For a fixed $j$, let $S_j^\text{true} = 1$, and $S_k^\text{true} = 0\,\forall k\neq j$.
    \item Draw simulated events $\lbrace x_t^{(k)}\rbrace = \lbrace(\vec{\nu}^{(k)}_t,\vec\theta^{(k)}_t)\rbrace$ from the PDF $P_j$,
        \begin{equation}
            P_j \equiv P(\vec\nu,\vec\theta | \lbrace S_j = 1, S_{k\neq j} = 0\rbrace)\,.
        \end{equation}
    \item Use the true events as input for the MC simulation of the detector, in order to obtain the detector
        events $\lbrace x^{(k)}_d\rbrace = \lbrace(\vec{\nu}^{(k)}_d,\vec\theta^{(k)}_d)\rbrace$.
    \item Determine the uncorrected observables $\lbrace \est{S_k} \rbrace$ based on the detector
        events $\lbrace x^{(k)}_d\rbrace$. For all intents and purposes, automation of this step is only feasible
        with to the method of moments, since it does not suffer form convergence problems.
    \item By virtue of \refeq{linear-assumption} and step 1, the $j$-th column of $M$ has been determined via
        \begin{equation}
            M_{ij} = \est{S_i}\,.
        \end{equation}
    \item Repeat the above steps for all values of $j$ to determine all columns of $M$.
\end{enumerate}

@@

We emphasize that this methods hinges on the assumption in \refeq{linear-assumption}; we thus suggest to explicitly
test the linearity. One way of testing involves repeating the previous procedure
with $S_j^\text{true} = 1/z$, $z > 0$. If the linearity assumption holds, one should find
\begin{equation}
    M_{ij}(z) = z \est{S_i}\,,
\end{equation}
for a fixed value of $j$.

@@

In order to study the performance of the proposed method, we simulate measurements for one uni-angular
and one tri-angular decay distribution for various numbers of events. As inputs we use the SM predictions
for the decays $B\to K\ell^+\ell^-$ and $B\to K^*(\to K\pi)\ell^+\ell^-$, respectively. We cover the
$q^2$ ranges $1\GeV^2$ to $6\GeV^2$, and $15\GeV^2$ to to the respective kinematic endpoint, with two
bin widths: $1\GeV^2$ and $0.5\GeV^2$. This is to ensure that a wide spectrum of possible values for
the angular observables.\\

Our finding can be summarized as follows:
\begin{itemize}
    \item In all studied cases we observed not a single bias in the pull distribution of the observable.
        All pull distributions can successfully be fitted to a gaussian distribution. As an example, we
        show the pull distributions for the observables $S_5$ and $S_7$ for a SM-like $B\to K^*\ell^+\ell^-$
        decays in \reffig{pulls}.
    \item We study the absolute uncertainty $\sigma_k(N)$ with respect to the angular observable $S_k$
        as a function of the number of simulated events $N$. As expected for a multivariate gaussian
        distribution, we find that the absolute uncertainty is well fitted by
        \begin{equation}
            \sigma_k(N) = \frac{\sigma_k(1)}{\sqrt{N}}
        \end{equation}
        with $\sigma_k(1) = \order{1}$, regardless of the size of $S_k$. The latter can best be shown
        for the uncertainties of two exampular observables. Taking again $S_5$ ($\simeq 27\%$) and $S_7$ ($\simeq 2\%$)
        for SM-like $B\to K^*\ell^+\ell^-$ decays, we show the absolute uncertainty in \reffig{errors}.
    \item For all toy studies, we produce plots of their Skewness and their Kurtosis as functions of
        $N$. We find in all cases that the distributions agree. 
    \item \todo{Plot Skewness and Kurtosis for $S_5$ and $S_7$ as functions of $N$.}
    \item \dots
\end{itemize}
We find that all of the above results show consistenly convergence of the distribution for the angular observables
towards a multivariate normal distribution.

\begin{figure}[t]
        \centering
            \includegraphics[width=0.45\textwidth]{figs/pull-Q2_5_6_S5_200.png}
            \includegraphics[width=0.45\textwidth]{figs/pull-Q2_5_6_S7_200.png}
        \caption{Pull distribution for the angular observables $S_5$ and $S_7$, extracted from $200$ simulated events of the decay $B\to K^*\ell^+\ell^-$. The red curve represents a fit to a gaussian distribution.}
        \label{fig:pulls}
\end{figure}

\begin{figure}[t]
        \centering
            \includegraphics[width=0.45\textwidth]{figs/error-Q2_5_6_S5.png}
            \includegraphics[width=0.45\textwidth]{figs/error-Q2_5_6_S7.png}
        \caption{Uncertainty of the angular observables $S_5$ and $S_7$, extracted from simulated events of the decay $B\to K^*\ell^+\ell^-$ and as a function of the number of simulated events $N$. The red curve represents a fit to the function $\sigma_k(N) = \frac{\sigma_k(1)}{\sqrt{N}}$.}
        \label{fig:errors}
\end{figure}
